Bayesian exponential random graph modeling of KWL Burial network

### 2020-10-10 analysis/paper/paper.Rmd
edit and fix the node number for pre-E

<!--burial study- network analysis-->
Burials provide valuable information about past human behavior that can reflect not only the personal life and status of the deceased, but also the expressions of their descendants through grave goods and arrangements [@Pearson1999]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization. A common way to study connections between individuals is social network analysis (SNA), which has been applied increasingly to archaeological studies [@Mills2017]. A network contains a set of actors and a set of relationships connecting these actors, which could be persons, groups, or organizations. In other words, actors are "nodes", in which nodes are connected by multiple relationships, or "edges", to other nodes. The basic unit for SNA analysis is dyad relationship, which is defined as a connection between two distinct actors [@Lee2016]. Relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relation in a given society [@Marin2011]. Using SNA, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the structure of the network [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]

@@

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures with a focus directly on possible ties between nodes by viewing them as random variables [@Robins2007]. In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measured separately by the presence of a tie without consideration of interactions among ties [@Morris2008]. By contrast, statistical models assumes social networks as dependent variables and allows a direct modeling for the network dependencies [@Snijders2011]. In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or the attributes of nodes/ties, called covariates. One can model network variables, such as dependencies, nodal attributes, and dyadic attributes, to generate edges between nodes based on a common experience for observed relationships. For example, common network dependencies include reciprocity that describes two actors' relationship, and transitivity that describes two actors' relationship via a shared third actor. Those variables, also called network statistics, are direct functions of the ties themselves where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. The advantage of networks specification under assumptions and contexts is useful to explain archaeological network in a given time. For instance, @Amati2019 demonstrated the use of ERGMs to reconstruct three different network for 15 sites in the Caribbean between AD 100 to 400 to explore the interaction mechanisms. Those models test the hypotheses for the relations between sites in terms of proximity, inter-cultural, and exchange network to explain the presence of hub sites. They found that the observed presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of the hypothesis of proximity or cultural homophily exclusively. However, they also point out some limitations of application, such as static outcome and sensitivity to missing data. In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits ERGMs applications. A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].

>>>>

<!--burial study- network analysis-->
Burials provide valuable information about past human behavior that can reflect not only the personal life and status of the deceased, but also the expressions of their descendants through grave goods and arrangements [@Pearson1999]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization. A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017]. Networks are represented by graphs that consists of a set of actors, such as people, groups, or organizations, and a set of links connecting them. Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes. The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011]. With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]

@@

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures by directly modeling the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that an observed network is randomly generated depending on the presence or absence of other edges. ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008]. In statistical models, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011]. For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby. Similarly, for network dependencies, reciprocity represents the likelihood between two actor's to be tied because of shared properties, and transitivity means two actors tend to be linked via a relationship with a third actor in common. We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. Network specification under assumptions and contexts is useful for exploring archaeological network. For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs. Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed. They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively. However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data. In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits the widely use of ERGMs. Here, we used ERGMs in a Bayesian framework that allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].
